---
output: 
  github_document:
    html_preview: true
bibliography: vignettes/lspline.bib
editor_options:
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `lspline`: Linear Splines with Convenient Parameterizations

<!-- badges: start -->
[![R-CMD-check](https://github.com/mbojan/lspline/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mbojan/lspline/actions/workflows/R-CMD-check.yaml)
[![rstudio mirror downloads](http://cranlogs.r-pkg.org/badges/lspline?color=2ED968)](http://cranlogs.r-pkg.org/)
[![cran version](http://www.r-pkg.org/badges/version/lspline)](https://cran.r-project.org/package=lspline)
<!-- badges: end -->


Linear splines with convenient parameterizations such that:

- coefficients are slopes of consecutive segments
- coefficients capture slope change at consecutive knots

Knot locations can be specified

- manually (`lspline()`)
- at breaks dividing the range of `x` into `q` equal-frequency intervals (`qlspline()`)
- at breaks dividing the range of `x` into `n` equal-width intervals (`elspline()`)

Inspired by Stata command `mkspline` and function `lspline()` from package **ares** [@r-ares]. As such, the implementation follows @greene2003econometric, chapter 7.5.2.



## Installation

From CRAN or development version using **remotes**:

```{r install, eval=FALSE}
remotes::install_github("mbojan/lspline", build_vignettes = TRUE)
```

## Examples

See package [homepage](https://mbojan.github.io/lspline/) and the [vignette](https://mbojan.github.io/lspline/articles/lspline.html).



```{r logo, include=FALSE, eval=FALSE}
# library(hexSticker)
library(tidyverse)
library(ggforce)
d1 <- tibble::tribble(
  ~x, ~y,
  0, 0,
  4, 1, # b = 1/4
  8, -1, # b = 1/2
  10, 2 # b = 3
) 
d2 <- tibble::tribble(
  ~x, ~y,
  4, 1,
  6, 1.5
)
d3 <- tibble::tribble(
  ~x, ~y,
  8, -1,
  10, -2,
)

arcs <- tibble::tribble(
  ~x0, ~y0, ~r, ~start, ~end,
  4, 1, 0.5, pi/2 - atan(1/4), pi/2 + atan(1/2),
  8, -1, 0.5, pi/2 - atan(3/2), pi/2 + atan(1/2)
)

d1 |>
  ggplot() +
  geom_line(aes(x=x, y=y), data = d1, linetype = "solid") +
  geom_line(aes(x=x, y=y), data = d2, linetype = "dotted") +
  geom_line(aes(x=x, y=y), data = d3, linetype = "dotted") +
  geom_arc(aes(x0 = x0, y0 = y0, r = r, start = start, end = end),
           data = arcs) +
  coord_fixed()


d <- tibble(
  y = c(0, 2, 1, 1.5),
  x = seq(along = y)
)

ggplot() +
  geom_line(aes(x=x, y=y), data = d, linetype = "solid") +
  coord_fixed()
  


d <- tibble::tribble(
  ~x, ~y,
  0, 0,
  2, 2,
  3, 3,
  5, 1,
  6, 0
) |>
  map(embed, 2) |>
  imap(~ structure(.x, dimnames=list(NULL, paste0(.y, c('end', ''))))) |>
  map(as_tibble) |>
  bind_cols() |>
  mutate(
    lt = rep(c("s", "d"), length = 4)
  )
  
d |>
  ggplot(aes(x=x, y=y, xend=xend, yend=yend)) +
  geom_segment(data = filter(d, lt == "s")) +
  scale_linetype_manual(
    breaks = c("s", "d"),
    values = c("solid", "dotted"),
    guide = "none"
  )

tibble::tribble(
  ~x, ~y, ~s,
  0, 0, "s",
  2, 2, "s",
  3, 3, "d",
  4, 4, "d"
) |>
  ggplot(aes(x = x, y = y, group = NULL)) +
  geom_line(aes(linetype = s))

```



## References
