---
output: 
  github_document:
    html_preview: true
bibliography: vignettes/lspline.bib
editor_options:
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `lspline`: Linear Splines with Convenient Parameterizations

<!-- badges: start -->
[![R-CMD-check](https://github.com/mbojan/lspline/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mbojan/lspline/actions/workflows/R-CMD-check.yaml)
[![rstudio mirror downloads](http://cranlogs.r-pkg.org/badges/lspline?color=2ED968)](http://cranlogs.r-pkg.org/)
[![cran version](http://www.r-pkg.org/badges/version/lspline)](https://cran.r-project.org/package=lspline)
<!-- badges: end -->


Linear splines with convenient parameterizations such that:

- coefficients are slopes of consecutive segments
- coefficients capture slope change at consecutive knots

Knot locations can be specified

- manually (`lspline()`)
- at breaks dividing the range of `x` into `q` equal-frequency intervals (`qlspline()`)
- at breaks dividing the range of `x` into `n` equal-width intervals (`elspline()`)

Inspired by Stata command `mkspline` and function `lspline()` from package **ares** [@r-ares]. As such, the implementation follows @greene2003econometric, chapter 7.5.2.



## Installation

From CRAN or development version using **remotes**:

```{r install, eval=FALSE}
remotes::install_github("mbojan/lspline", build_vignettes = TRUE)
```

## Examples

See package [homepage](https://mbojan.github.io/lspline/) and the [vignette](https://mbojan.github.io/lspline/articles/lspline.html).

## References
