---
title: "Fitting linear splines"
output: 
  rmarkdown::html_vignette:
    toc: true
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Fitting linear splines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  cache=FALSE,
  collapse=TRUE
)
```


Here are some examples of using `lspline()`, `qlspline()`, and `elspline()`.


# Generate example data

We will use the following artificial data with knots at `x=5` and `x=10`

```{r data}
set.seed(666)
n <- 200
d <- data.frame(
  x = scales::rescale(rchisq(n, 6), c(0, 20))
)
d$interval <- findInterval(d$x, c(5, 10), rightmost.closed = TRUE) + 1
d$slope <- c(2, -3, 0)[d$interval]
d$intercept <- c(0, 25, -5)[d$interval]
d$y <- with(d, intercept + slope * x + rnorm(n, 0, 1))
```

Plotting `y` against `x`:

```{r show_data}
library(ggplot2)
fig <- ggplot(d, aes(x=x, y=y)) + 
  geom_point(aes(shape=as.character(slope))) +
  scale_shape_discrete(name="Slope") +
  theme_bw()
fig
```

The slopes of the consecutive segments are 2, -3, and 0.




## Setting knot locations manually

We can parametrize the spline with slopes of individual segments (default `marginal=FALSE`):

```{r lspline_nonmarginal}
library(lspline)
m1 <- lm(y ~ lspline(x, c(5, 10)), data=d)
knitr::kable(broom::tidy(m1))
```

Or parametrize with coeficients measuring change in slope (with `marginal=TRUE`):

```{r rspline_marginal, echo=1:2}
m2 <- lm(y ~ lspline(x, c(5,10), marginal=TRUE), data=d)
knitr::kable(broom::tidy(m2))
k <- coef(m2)
nam <- names(k)
```

The coefficients are

- ``r nam[2]`` - the slope of the first segment
- ``r nam[3]`` - the change in slope at knot $x=5$; it is changing from 2 to -3, so by -5
- ``r nam[4]`` - tha change in slope at knot $x=10$; it is changing from -3 to 0, so by 3

The two parametrisations (obviously) give identical predicted values:

```{r same-fitted}
all.equal( fitted(m1), fitted(m2) )
```

graphically

```{r lspline_fitted}
fig +
  geom_smooth(method="lm", formula=formula(m1), se=FALSE) +
  geom_vline(xintercept = c(5, 10), linetype=2)
```




## Knots at `n` equal-length intervals

Function `elspline()` sets the knots at points dividing the range of `x` into `n` equal length intervals.

```{r elspline}
m3 <- lm(y ~ elspline(x, 3), data=d)
knitr::kable(broom::tidy(m3))
```

Graphically

```{r elspline-fitted}
fig +
  geom_smooth(aes(group=1), method="lm", formula=formula(m3), se=FALSE, n=200)
```




## Knots at `q`uantiles of `x`

Function `qlspline()` sets the knots at points dividing the range of `x` into `q` equal-frequency intervals.

```{r qlspline}
m4 <- lm(y ~ qlspline(x, 4), data=d)
knitr::kable(broom::tidy(m4))
```

Graphically

```{r qlspline-fitted}
fig +
  geom_smooth(method="lm", formula=formula(m4), se=FALSE, n=200)
```
